---
title: "Data merge"
author: "NEST coreDev"
date: "2022-05-11"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data merge}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Combining datasets is a crucial step when using modules with more than one dataset. In the context of `teal`, we use 
the term "merge" to combine datasets where two functions are offered `data_merge_module` and `data_merge_srv`. Depending on the specific scenario, one or the other shall be used. 

When no processing of the `data_extract` list is required, the `data_merge_module` function is used to read the data and the `data_extract_spec`'s list and apply the merging. It is a wrapper that combines `data_extract_multiple_srv()` and `data_merge_srv()` see below for more details.
With additional processing of the `data_extract` list input, `data_merge_srv()` can be combined with `data_extract_multiple_srv()` or `data_extract_srv()` to shape the `selector_list` input. 

In the coming sections, we will show examples of both scenarios.

### `data_merge_module`

With `data_merge_module` solely, all you would need is a list of `data_extract_spec` objects for the `data_extract` argument and a `FilteredData` object for the `datasets` argument.

#### App code

```{r eval=FALSE}
library(teal.transform)
library(shiny)

adsl_extract <- teal.transform::data_extract_spec(
  dataname = "ADSL",
  select = select_spec(
    label = "Select variable:",
    choices = c("AGE", "BMRKR1"),
    selected = "AGE",
    multiple = TRUE,
    fixed = FALSE
  )
)

adtte_extract <- teal.transform::data_extract_spec(
  dataname = "ADTTE",
  select = select_spec(
    choices = c("AVAL", "ASEQ"),
    selected = "AVAL",
    multiple = TRUE,
    fixed = FALSE
  )
)

data_extracts <- list(adsl_extract = adsl_extract, adtte_extract = adtte_extract)

merge_ui <- function(id, data_extracts) {
  ns <- NS(id)
  teal.widgets::standard_layout(
    output = teal.widgets::white_small_well(
      verbatimTextOutput(ns("expr")),
      dataTableOutput(ns("data"))
    ),
    encoding = div(
      teal.transform::data_extract_ui(
        ns("adsl_extract"), # must correspond with data_extracts list names
        label = "ADSL extract",
        data_extracts[[1]]
      ),
      teal.transform::data_extract_ui(
        ns("adtte_extract"), # must correspond with data_extracts list names
        label = "ADTTE extract",
        data_extracts[[2]]
      )
    )
  )
}

merge_module <- function(id, datasets, data_extracts) {
  moduleServer(id, function(input, output, session) {
    merged_data <- teal.transform::data_merge_module(
      data_extract = data_extracts,
      datasets = datasets,
      merge_function = "dplyr::left_join"
    )
    output$expr <- renderText(merged_data()$expr)
    output$data <- renderDataTable(merged_data()$data())
  })
}

sample_filtered_data <- function() {
  # create TealData
  adsl <- teal.data::cdisc_dataset("ADSL", scda::synthetic_cdisc_data("latest")$adsl)
  adtte <- teal.data::cdisc_dataset("ADTTE", scda::synthetic_cdisc_data("latest")$adtte)
  data <- teal.data::cdisc_data(adsl, adtte)

  # convert TealData to FilteredData
  datasets <- teal.slice:::filtered_data_new(data)
  teal.slice:::filtered_data_set(data, datasets)
  datasets
}

datasets <- sample_filtered_data()
```

#### Shiny app

```{r eval=FALSE}
shinyApp(
  ui = fluidPage(merge_ui("data_merge", data_extracts)),
  server = function(input, output, session) {
    merge_module("data_merge", datasets, data_extracts)
  }
)
```

### `data_extract_multiple_srv` + `data_merge_srv`

In the scenario above, if the user deselects the `ADTTE` variable, the merging between `ADTTE` and `ADSL` would still take place even though `ADTTE` is not used or needed here. Here, the developer might update the `selector_list` input in a reactive manner so that it gets updated based on conditions set by the developer. Below, we reuse the input from above and we update the `app` server so that the `adtte_extract` is removed from the `selector_list` input when no `ADTTE` variable is selected and the `reactive_selector_list` is passed to `data_merge_srv`:

```{r eval=FALSE}
merge_module <- function(id, datasets, data_extracts) {
  moduleServer(id, function(input, output, session) {
    selector_list <- teal.transform::data_extract_multiple_srv(data_extracts, datasets)
    reactive_selector_list <- reactive({
      if (is.null(selector_list()$adtte_extract) || length(selector_list()$adtte_extract()$select) == 0) {
        selector_list()[names(selector_list()) != "adtte_extract"]
      } else {
        selector_list()
      }
    })

    merged_data <- teal.transform::data_merge_srv(
      selector_list = reactive_selector_list,
      datasets = datasets,
      merge_function = "dplyr::left_join"
    )
    output$expr <- renderText(merged_data()$expr)
    output$data <- renderDataTable(merged_data()$data())
  })
}
```

#### Shiny app

```{r eval=FALSE}
shinyApp(
  ui = fluidPage(merge_ui("data_merge", data_extracts)),
  server = function(input, output, session) {
    merge_module("data_merge", datasets, data_extracts)
  }
)
```

`data_merge_module` is replaced here with three parts:

1) `selector_list`: output of `data_extract_multiple_srv` which loops over the list of data_extract given and runs `data_extract_srv` for each one returning a list of reactive objects.
2) `reactive_selector_list`: intermediate reactive list updating `selector_list` content
3) `merged_data`: output of `data_merge_srv` using `reactive_selector_list` as input

Both merge functions, `data_merge_srv` and `data_merge_module`, return a reactive object which contains a list of the following elements:

-`data`: the merged dataset after filtering and reshaping containing selected columns

-`expr`: code needed to replicate merged dataset

-`chunks`: chunks R6 object (see `teal.code`)

-`columns_source`: list of columns selected per selector

-`keys`: the keys of the merged dataset

-`filter_info`: filters that are applied on the data

These elements can be further used inside the server to retrieve and use information about the selections, data, filters ...

## Merging of non `CDISC` datasets

General datasets do not share the same relationships as `CDISC` datasets thus these relationships must be specified by the `join_keys` functions. For more information, please refer to the `Join Keys` [vignette](https://insightsengineering.github.io/teal/join-keys).

The data merge module respects the relationships given by the user and in the case of multiple datasets to merge, the order is specified by the order of elements in the `data_extract` argument of the `data_merge_module` function. Merging groups of datasets with complex relationships can quickly become challenging to specify so please take extra care when setting this up.
