---
title: "Data extract"
author: "NEST coreDev"
date: "2022-05-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data extract}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are times when an app developer wants to showcase more than just one fixed slice of their dataset in their
custom module. Relinquishing control of the application to a user demands the developer gives their users a degree
of freedom. In case of analyzing data, `teal` allows app developers to open up their applications to users, letting them
decide exactly what app data to analyze in the module.

A lot of `teal` modules use `data_extract_spec` objects and modules to tackle user input. You can find many examples in
e.g. `teal.modules.general` and `teal.modules.clinical`.


### `data_extract_spec`

`data_extract_spec`'s task is two-fold: create a UI component in a `shiny` application and pass the user input from the
 UI to the module itself. Having that formulated, let's have a look at how it supports both its responsibilities.

#### Example module

In order to showcase different initialization options of `data_extract_spec`, first we define a `shiny` module which
uses `data_extract_ui` and `data_extract_srv` designed to handle `data_extract_spec` objects. The module creates a UI
component for single `data_extract_spec` and prints list of values returned from `data_extract_srv` module. Please see
package documentation for more information about `data_extract_ui` and `data_extract_srv`.

```{r}
library(teal.transform)
library(shiny)

extract_ui <- function(id, data_extract) {
  ns <- NS(id)
  teal.widgets::standard_layout(
    output = teal.widgets::white_small_well(verbatimTextOutput(ns("output"))),
    encoding = data_extract_ui(ns("data_extract"), label = "variable", data_extract)
  )
}

extract_srv <- function(id, datasets, data_extract) {
  moduleServer(id, function(input, output, session) {
    reactive_extract_input <- data_extract_srv("data_extract", datasets, data_extract)
    s <- reactive({
      format_data_extract(reactive_extract_input())
    })
    output$output <- renderPrint({
      cat(s())
    })
  })
}
```


#### Example data

`teal.transform` functions depend on a `FilteredData` object from the `teal.slice` package. Normally, `FilteredData` is created
automatically by `teal::init`, but for example purposes we define a wrapper function to initialize the necessary object.

```{r}
sample_filtered_data <- function() {
  # create TealData
  adsl <- teal.data::cdisc_dataset("ADSL", scda::synthetic_cdisc_data("latest")$adsl)
  adtte <- teal.data::cdisc_dataset("ADTTE", scda::synthetic_cdisc_data("latest")$adtte)
  data <- teal.data::cdisc_data(adsl, adtte)

  # covert TealData to FilteredData
  datasets <- teal.slice:::filtered_data_new(data)
  teal.slice:::filtered_data_set(data, datasets)
  datasets
}

datasets <- sample_filtered_data()
```

Consider the following example, where we create two UI elements, one to filter on a specific level from `SEX` variable,
and a second one to select a variable from `c("BMRKR1", "AGE")`. `data_extract_spec` object is handed over to the shiny
app and gives instructions to generate UI components.

```{r}
simple_des <- data_extract_spec(
  dataname = "ADSL",
  filter = filter_spec(vars = "SEX", choices = c("F", "M")),
  select = select_spec(choices = c("BMRKR1", "AGE"))
)
```

#### Shiny app

Finally, we pass `extract_ui` to the `ui` of the `shinyApp` and we use `extract_srv` in the server function of the `shinyApp`:

```{r eval=FALSE}
shinyApp(
  ui = fluidPage(extract_ui("data_extract", simple_des)),
  server = function(input, output, session) {
    extract_srv("data_extract", datasets, simple_des)
  }
)
```
