---
title: "Data Extract"
author: "NEST CoreDev"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Extract}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are times when an app developer wants to showcase more than just one fixed slice of their dataset in their
custom module. Relinquishing control of the application to a user demands the developer gives their users a degree
of freedom. In case of analyzing data, `teal` allows app developers to open up their applications to users, letting them
decide exactly what app data to analyze in the module.

A lot of `teal` modules use `data_extract_spec` objects and modules to tackle user input. You can find many examples in
e.g. `teal.modules.general` and `teal.modules.clinical`.


### `data_extract_spec`

`data_extract_spec`'s task is two-fold: create a UI component in a `shiny` application and pass the user input from the
 UI to the module itself. Having that formulated, let's have a look at how it supports both its responsibilities.

#### Example module

In order to showcase different initialization options of `data_extract_spec`, first we define a `shiny` module which
uses `data_extract_ui` and `data_extract_srv` designed to handle `data_extract_spec` objects. The module creates a UI
component for a single `data_extract_spec` and prints a list of values returned from `data_extract_srv` module. Please see
package documentation for more information about `data_extract_ui` and `data_extract_srv`.

```{r}
library(teal.transform)
library(shiny)

extract_ui <- function(id, data_extract) {
  ns <- NS(id)
  teal.widgets::standard_layout(
    output = teal.widgets::white_small_well(verbatimTextOutput(ns("output"))),
    encoding = data_extract_ui(ns("data_extract"), label = "variable", data_extract)
  )
}

extract_srv <- function(id, datasets, data_extract, join_keys) {
  moduleServer(id, function(input, output, session) {
    reactive_extract_input <- data_extract_srv("data_extract", datasets, data_extract, join_keys)
    s <- reactive({
      format_data_extract(reactive_extract_input())
    })
    output$output <- renderPrint({
      cat(s())
    })
  })
}
```


#### Example data

`data_extract_srv` module depends on either a list of reactive or non-reactive `data.frame` objects. Here, we show the
usage of a list of `data.frame` objects as input to `datasets`
where a list of necessary join keys per `data.frame` object is required:

```{r}
# Define data.frame objects
ADSL <- rADSL # nolint
ADTTE <- rADTTE # nolint

# create a list of data.frame objects
datasets <- list(ADSL = ADSL, ADTTE = ADTTE)

# create join_keys
join_keys <- teal.data::join_keys(
  teal.data::join_key("ADSL", "ADSL", c("STUDYID", "USUBJID")),
  teal.data::join_key("ADSL", "ADTTE", c("STUDYID", "USUBJID")),
  teal.data::join_key("ADTTE", "ADTTE", c("STUDYID", "USUBJID", "PARAMCD"))
)
```

Consider the following example, where we create two UI elements, one to filter on a specific level from `SEX` variable,
and a second one to select a variable from `c("BMRKR1", "AGE")`. `data_extract_spec` object is handed over to the shiny
app and gives instructions to generate UI components.

```{r}
simple_des <- data_extract_spec(
  dataname = "ADSL",
  filter = filter_spec(vars = "SEX", choices = c("F", "M")),
  select = select_spec(choices = c("BMRKR1", "AGE"))
)
```

#### Shiny app

Finally, we pass `extract_ui` to the `ui` of the `shinyApp` and we use `extract_srv` in the server function of the `shinyApp`:

```{r eval=FALSE}
shinyApp(
  ui = fluidPage(extract_ui("data_extract", simple_des)),
  server = function(input, output, session) {
    extract_srv("data_extract", datasets, simple_des, join_keys)
  }
)
```
