---
title: "Data Merge"
author: "NEST CoreDev"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Merge}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Joining datasets is an essential step when working with relational datasets.

To support this, two functions are provided depending on how to process the `data_extract_spec` object:

1. `merge_expression_module` can be used when there is no need to process the list of `data_extract_spec`.
This function reads the data and the list of `data_extract_spec` objects and applies the merging.
Essentially, it serves as a wrapper that combines `data_extract_multiple_srv()` and `merge_expression_srv()`.
2. `merge_expression_srv` and `data_extract_multiple_srv` can be used in scenarios where additional processing of the list of `data_extract_spec` is necessary or `data_extract_srv()` to customize the `selector_list` input.

The following sections provide examples for both scenarios.

### `merge_expression_module`

Using `merge_expression_module` alone requires a list of `data_extract_spec` objects for the `data_extract` argument,
a list of reactive or non-reactive `data.frame` objects, and a list of join keys corresponding to each `data.frame` object.

#### Step 1/5 - Preparing the Data

```{r library}
library(teal.transform)
library(teal.data)
library(shiny)

# Define data.frame objects
ADSL <- teal.data::rADSL
ADTTE <- teal.data::rADTTE

# create a list of reactive data.frame objects
datasets <- list(
  ADSL = reactive(ADSL),
  ADTTE = reactive(ADTTE)
)

# create  join_keys
join_keys <- join_keys(
  join_key("ADSL", "ADSL", c("STUDYID", "USUBJID")),
  join_key("ADSL", "ADTTE", c("STUDYID", "USUBJID")),
  join_key("ADTTE", "ADTTE", c("STUDYID", "USUBJID", "PARAMCD"))
)
```


#### Step 2/5 - Creating the Data Extracts

```{r data_extract_spec}
adsl_extract <- data_extract_spec(
  dataname = "ADSL",
  select = select_spec(
    label = "Select variable:",
    choices = c("AGE", "BMRKR1"),
    selected = "AGE",
    multiple = TRUE,
    fixed = FALSE
  )
)

adtte_extract <- data_extract_spec(
  dataname = "ADTTE",
  select = select_spec(
    choices = c("AVAL", "ASEQ"),
    selected = "AVAL",
    multiple = TRUE,
    fixed = FALSE
  )
)

data_extracts <- list(adsl_extract = adsl_extract, adtte_extract = adtte_extract)
```

#### Step 3/5 - Creating the UI

```{r merge_ui}
merge_ui <- function(id, data_extracts) {
  ns <- NS(id)
  sidebarLayout(
    sidebarPanel(
      h3("Encoding"),
      tags$div(
        data_extract_ui(
          ns("adsl_extract"), # must correspond with data_extracts list names
          label = "ADSL extract",
          data_extracts[[1]]
        ),
        data_extract_ui(
          ns("adtte_extract"), # must correspond with data_extracts list names
          label = "ADTTE extract",
          data_extracts[[2]]
        )
      )
    ),
    mainPanel(
      h3("Output"),
      verbatimTextOutput(ns("expr")),
      dataTableOutput(ns("data"))
    )
  )
}
```

#### Step 4/5 - Creating the Server Logic

```{r merge_srv}
merge_srv <- function(id, datasets, data_extracts, join_keys) {
  moduleServer(id, function(input, output, session) {
    merged_data <- merge_expression_module(
      data_extract = data_extracts,
      datasets = datasets,
      join_keys = join_keys,
      merge_function = "dplyr::left_join"
    )

    ANL <- reactive({
      data_list <- lapply(datasets, function(ds) ds())
      eval(envir = list2env(data_list), expr = as.expression(merged_data()$expr))
    })
    output$expr <- renderText(paste(merged_data()$expr, collapse = "\n"))
    output$data <- renderDataTable(ANL())
  })
}
```

#### Step 5/5 - Creating the `shiny` App

```{r shinyapp, eval=FALSE}
shinyApp(
  ui = fluidPage(merge_ui("data_merge", data_extracts)),
  server = function(input, output, session) {
    merge_srv("data_merge", datasets, data_extracts, join_keys)
  }
)
```

```{r shinylive_url, echo = FALSE, results = 'asis', eval = requireNamespace("roxy.shinylive", quietly = TRUE)}
code <- paste0(c(
  knitr::knit_code$get("library"),
  knitr::knit_code$get("data_extract_spec"),
  knitr::knit_code$get("merge_ui"),
  knitr::knit_code$get("merge_srv"),
  knitr::knit_code$get("shinyapp")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
cat(sprintf("[Open in Shinylive](%s)\n\n", url))
```

```{r shinylive_iframe, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
knitr::include_url(url, height = "800px")
```

<hr>

### `data_extract_multiple_srv` + `merge_expression_srv`

In the scenario above, if the user deselects the `ADTTE` variable, the merging between `ADTTE` and `ADSL` would still occur, even though `ADTTE` is not used or needed.
Here, the developer might update the `selector_list` input in a reactive manner so that it gets updated based on conditions set by the developer.
Below, we reuse the input from above and update the app server so that the `adtte_extract` is removed from the selector_list input when no `ADTTE` variable is selected.
The `reactive_selector_list` is then passed to `merge_expression_srv`:

#### Modifying the Server Logic

```{r merge_srv2}
merge_srv <- function(id, datasets, data_extracts, join_keys) {
  moduleServer(id, function(input, output, session) {
    selector_list <- data_extract_multiple_srv(data_extracts, datasets, join_keys)
    reactive_selector_list <- reactive({
      if (is.null(selector_list()$adtte_extract) || length(selector_list()$adtte_extract()$select) == 0) {
        selector_list()[names(selector_list()) != "adtte_extract"]
      } else {
        selector_list()
      }
    })

    merged_data <- merge_expression_srv(
      selector_list = reactive_selector_list,
      datasets = datasets,
      join_keys = join_keys,
      merge_function = "dplyr::left_join"
    )

    ANL <- reactive({
      data_list <- lapply(datasets, function(ds) ds())
      eval(envir = list2env(data_list), expr = as.expression(merged_data()$expr))
    })
    output$expr <- renderText(paste(merged_data()$expr, collapse = "\n"))
    output$data <- renderDataTable(ANL())
  })
}
```

#### Updating the `shiny` app

```{r shinyapp2, eval=FALSE}
shinyApp(
  ui = fluidPage(merge_ui("data_merge", data_extracts)),
  server = function(input, output, session) {
    merge_srv("data_merge", datasets, data_extracts, join_keys)
  }
)
```

```{r shinylive_url2, echo = FALSE, results = 'asis', eval = requireNamespace("roxy.shinylive", quietly = TRUE)}
code <- paste0(c(
  knitr::knit_code$get("library"),
  knitr::knit_code$get("data_extract_spec"),
  knitr::knit_code$get("merge_ui"),
  knitr::knit_code$get("merge_srv2"),
  knitr::knit_code$get("shinyapp2")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
cat(sprintf("[Open in Shinylive](%s)\n\n", url))
```

```{r shinylive_iframe2, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
knitr::include_url(url, height = "800px")
```

`merge_expression_module` is replaced here with three parts:

1) `selector_list`: the output of `data_extract_multiple_srv`, which loops over the list of `data_extract` given and runs `data_extract_srv` for each one, returning a list of reactive objects.
2) `reactive_selector_list`: an intermediate reactive list updating `selector_list` content.
3) `merged_data`: the output of `merge_expression_srv` using `reactive_selector_list` as input.

### Output from merging

Both merge functions, `merge_expression_srv` and `merge_expression_module`, return a reactive object which contains a list of the following elements:

-  `expr`: code needed to replicate merged dataset.
-  `columns_source`: list of columns selected per selector.
-  `keys`: the keys of the merged dataset.
-  `filter_info`: filters that are applied on the data.

These elements can be further used inside the server to retrieve and use information about the selections, data, filters, etc.

## Merging of non `CDISC` datasets

General datasets do not have the same relationships as `CDISC` datasets, so these relationships must be specified using the `join_keys` functions.
For more information, please refer to the `Join Keys` [vignette](https://insightsengineering.github.io/teal.data/latest-tag/articles/join-keys).
The data merge module respects the relationships given by the user.
In the case of multiple datasets to merge, the order is specified by the order of elements in the `data_extract` argument of the `merge_expression_module` function.
Merging groups of datasets with complex relationships can quickly become challenging to specify so please take extra care when setting this up.
